---
title: "Use of Generative AI to Identify 4 Helmet Status Among Patients With Micromobility-Related Injuries From Unstructured Clinical Notes"
author: "Kathryn Burford"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library("tidyverse")
library("dplyr")
library("data.table")
library("readr")
library("survey")
library("janitor")
library("naniar")
library("gtsummary")
library("readxl")
library("ggplot2")
library("hexbin")
library("ggsurvey")
library("irr")
library("psych")
library(irr)
library(rlang)
library(reshape2)
library(epiR)
library(skimr)
library(DescTools)
library(purrr)
library(caret)
library(purrr)
library(irr)
library(stringr)
```

#read dirty neiss data
```{r}
neiss2019 <- read_excel("data/neiss_2019_2022/neiss2019.xlsx")
neiss2020 <- read_excel("data/neiss_2019_2022/neiss2020.xlsx")
neiss2021 <- read_excel("data/neiss_2019_2022/neiss2021.xlsx")
neiss2022 <- read_excel("data/neiss_2019_2022/neiss2022.xlsx")
```

#merge dataframes
```{r}
# Add a new column to represent the original dataframe
neiss2019 <- neiss2019 %>%
  mutate(year = 2019)

neiss2020 <- neiss2020 %>%
  mutate(year = 2020)

neiss2021 <- neiss2021 %>%
  mutate(year = 2021)

neiss2022 <- neiss2022 %>%
  mutate(year = 2022)

# Append the dataframes
df_dirty <- bind_rows(neiss2019, neiss2020, neiss2021, neiss2022)
df <- bind_rows(neiss2019, neiss2020, neiss2021, neiss2022)

```


#clean data 

```{r}
# tidy column names and select variables of interest
df <- df %>%
  janitor::clean_names() %>%
  select(cpsc_case_number, year, product_1, product_2, product_3, narrative_1)

```

##helmet code
```{r}

# HELMET 
# first pull -HELMET and - HELMET 
df$helmet_negative1 <- ifelse(str_detect(df$narrative_1, "\\-\\s?HELMET") , 1, 0)

# HELMET = NEGATIVE-----
# create negative helmet list- 
helmet_neg <- c("NO HELMET", "WITHOUT A HELMET", "WITHOUT HELMET", "NEGATIVE FOR HELMET", "NEGATIVE FOR A HELMET", 
"NOT HELMETED", "-HELMET", "- HELMET", "NOT WEARING HELMET", "NOT WEARING A HELMET", "UNHELMETED", 
"W/O HELMET", "DID NOT HAVE ON A HELMET", "NO WEARING HELMET", "DID NOT HAVE A HELMET", 
"DID NOT HAVE HELMET", "W/O A HELMET", "WITHOUT WEARING HELMET", "WITHOUT WEARING A HELMET", 
"REMOVED HELMET", "DENIED HELMET", "DENIED USE OF HELMET", "DENIED USE OF A HELMET", "HELMETLESS", 
"W/OUT HELMET", "DENIES HELMET", "WITH OUT HELMET", "WITH OUT A HELMET", "NOT WEAR A HELMET", 
"NOT WEAR HELMET", "WO A HELMET", "-HELMET", "NEGATIVE HELMET", "WIHTOUT A HELMET", "NOT WEARIING HELMET", "NOT WEARINNG A HELMET", "NOT WEARIG HELMET")


df <- df %>%
  mutate(helmet_negative2 = ifelse(Reduce(`|`, lapply(helmet_neg, str_detect, string = narrative_1)), 1, 0))

# create variable to combine the two positive variables and remove old variables
df$helmet_negative <- ifelse(df$helmet_negative1 == 1 | df$helmet_negative2 == 1, 1, 0)
df<- select(df, -c(helmet_negative1, helmet_negative2))


# spot check helment_negative = 0: these should all either be yes helmets or unclear
# df %>%
# dplyr::filter(helmet_negative==0) %>% View()

## HELMET = UNSURE-----

# first pull ?HELMET and ? HELMET 
df$helmet_unknown1 <- ifelse(str_detect(df$narrative_1, "\\?\\s?HELMET") , 1, 0) 
#first pull HELMET? 
df$helmet_unknown1.1 <- ifelse(str_detect(df$narrative_1, "HELMET\\?") , 1, 0)

# then pull all other unknown helmet phrases
# create unknown helmet list
helmet_unk <- c("UNSURE IF HELMET", "UNSURE IF HELMETED", "UNSURE IF PT WEARING HELMET", 
"UNSURE IF PT WEARING A HELMET", "UNSURE IF PT WAS WEARING HELMET", 
"UNSURE IF PT WAS WEARING A HELMET", "UNKNOWN IF HELMET", "HELMET UNKNOWN", 
"UNKNOWN HELMET", "NO MENTION OF HELMET", "UNK HELMET", "HELMET UNK", 
"HELMET NS", "NS HELMET")

#create variable indicating presence of one of the phrases
df <- df %>%
  mutate(helmet_unknown2 = ifelse(Reduce(`|`, lapply(helmet_unk, str_detect, string = narrative_1)), 1, 0))

#create new variable to combine the two unknowns and remove old variables
df$helmet_unknown <- ifelse(df$helmet_unknown1 == 1 | df$helmet_unknown1.1 ==1 | df$helmet_unknown2 == 1, 1, 0)
df <- select(df, -c(helmet_unknown1,helmet_unknown1.1, helmet_unknown2))

# spot check helmet_negative = 0 and helmet_unknown =0: these should all be yes or some weird wording
# df_helmet %>%
#  dplyr::filter(helmet_unknown==0 & helmet_negative==0) %>% View()

## HELMET = POSITIVE----
# first pull +HELMET and + HELMET (because of regex things)
df$helmet_positive1 <- ifelse(str_detect(df$narrative_1, "\\+\\s?HELMET") , 1, 0)
# first pull + FUll HELMET and +FULL HELMET (because of regex things)
df$helmet_positive1.1 <- ifelse(str_detect(df$narrative_1, "\\+\\s?FULL HELMET") , 1, 0)
# first pull + FUll BIKE HELMET and +FULL BIKE HELMET (because of regex things)
df$helmet_positive1.2 <- ifelse(str_detect(df$narrative_1, "\\+\\s?FULL BIKE HELMET") , 1, 0)
# first pull + BIKE HELMET and +BIKE HELMET (because of regex things)
df$helmet_positive1.3 <- ifelse(str_detect(df$narrative_1, "\\+\\s?BIKE HELMET") , 1, 0)
# first pull & HELMET (because of regex things)
df$helmet_positive1.4 <- ifelse(str_detect(df$narrative_1, "\\&\\s?HELMET") , 1, 0)
# first pull + HELMEMT (because of regex things)
df$helmet_positive1.5 <- ifelse(str_detect(df$narrative_1, "\\+\\s?HELMEMT") , 1, 0)

# create positive list and pull all records matching
helmet_pos <- c("WITH A HELMET", "WITH HELMET", "POSITIVE FOR HELMET", "POSITIVE FOR A HELMET", "HELMETED", 
"WEARING HELMET", "WEARING A HELMET", "W/ HELMET", "W HELMET", "WORE HELMET", "W/HELMET", 
"W/ A HELMET", "HAD ON A HELMET", "HAD ON HELMET", "HAD HELMET", "HAD A HELMET", "HIS HELMET", 
"HER HELMET", "PT HELMET", "PTS HELMET", "WAS WEARING A HELMET", ",HELMET,", "CRACKED HELMET", 
"CRACKING HELMET", "HELMET CRACKED", "HELMET ON", "BROKE HELMET", "HELMET BROKE", "BREAKING HELMET", 
"POSITIVE HELMET", "WORE HELMET", "WEARING A BIKE HELMET", "WEARING A FULL FACE HELMET", 
"WEARING FULL FACE HELMET", "INCLUDING HELMET", " FULL HELMET", "PT'S HELMET", 
"HELMET FELL", "SMASHED HELMET", "PLUS HELMET", "HELMET WAS FRACTURED", "HELMET WAS BROKEN", 
 "WEARING BIKE HELMET", "ENDORSES HELMET", "HELMET WENT OFF", 
"FULL FACE MASK HELMET", "& HELMET", "HELMET WAS REPORTED TO BE CRACKED", "HELMETD", "WHELMET")

df <- df %>%
  mutate(helmet_positive2 = ifelse(Reduce(`|`, lapply(helmet_pos, str_detect, string = narrative_1)), 1, 0))

# create variable to combine the two positive variables and remove old variables
df$helmet_positive3 <- ifelse(df$helmet_positive1 == 1 | df$helmet_positive1.1 == 1 | df$helmet_positive1.2  == 1 | df$helmet_positive1.3 == 1| df$helmet_positive1.4 == 1|  df$helmet_positive1.5 == 1|df$helmet_positive2 == 1, 1, 0)
df<- select(df, -c(helmet_positive1, helmet_positive1.1,helmet_positive1.2,helmet_positive1.3, helmet_positive1.4,helmet_positive1.5, helmet_positive2))

# subset only the positives that have not already been labeled as negative or unknown
df$cant_be_pos <- ifelse(df$helmet_negative == 1 | df$helmet_unknown == 1, 1, 0)
df$helmet_positive <- ifelse(df$helmet_positive3 == 1 & df$cant_be_pos == 0, 1, 0)

df<- select(df, -c(helmet_positive3, cant_be_pos))

nrow(df) - (sum(df$helmet_negative==1) + 
  sum(df$helmet_unknown==1) + sum(df$helmet_positive==1))

# Create a new variable 'helmet_unknown_all' for all of the unknowns (includes the "helmet_unknown" variable)
df$helmet_unknown_all <- ifelse(df$helmet_positive != 1 & df$helmet_negative != 1, 1, 0)


```

## 2019 micromobility variables

powered scooter, hoverboard, e-bike, bikes
```{r}

# Process 2019 separately bc product codes changed
# to get correct ebike, bike, scooter product code for 2019
# in 2020 discontinued 1329 and 5042 (scooter, powered) replaced with 5022, 5023 (unpowered), 5024
df_19 <- df %>%
  filter(year %in% c("2019"))

df_20_21_22 <- df %>%
  filter(year %in% c("2020", "2021", "2022"))

#note: all coded as 0=no, 1=yes

#POWERED SCOOTER (5042)
# scooter code
df_19$Code5042<-0
df_19$Code5042[df_19$product_1==5042 | df_19$product_2==5042 | df_19$product_3==5042]<-1

# scooter indicator
scooter1<-grep(c("scooter"), df_19$narrative_1, ignore.case = TRUE)
length(scooter1) 
df_19$scooter1<-0
df_19$scooter1[scooter1]<-1

df_19$scooter<-0
df_19$scooter[df_19$scooter1==1 & df_19$Code5042==1]<-1

# remove interim variables
df_19<-df_19[,!names(df_19) %in% c("scooter1")]

#  powered skateboard indicator
skateboard1<-grep(c("skate|skate board|skateboard"),df_19$narrative_1, ignore.case = TRUE)
length(skateboard1) 
df_19$skateboard1<-0
df_19$skateboard1[skateboard1]<-1

df_19$skateboard<-0
df_19$skateboard[df_19$skateboard1==1 & df_19$Code5042==1]<-1

# remove interim variables
df_19<-df_19[,!names(df_19) %in% c("skateboard1")]

#HOVERBOARD - 
# hoverboard variable
hover1<-grep(c("hoverboard|hooverboard|hover board|hoover board|hoverbaord|hooverbaord|hover boad|hoverboad,|hoverbrd"), df_19$narrative_1, ignore.case = TRUE)
length(hover1) 
df_19$hover1<-0
df_19$hover1[hover1]<-1

df_19$hover<-0
df_19$hover[df_19$hover1==1 & df_19$Code5042==1]<-1

# remove interim variables
df_19<-df_19[,!names(df_19) %in% c("hover1")]

# powered bike code
df_19$Codes5035.3215<-0
df_19$Codes5035.3215[df_19$product_1==5035 | df_19$product_2==5035 | df_19$product_3==5035| df_19$product_1==3215 | df_19$product_2==3215 |df_19$product_3==3215]<-1

# E-BIKE (5035, 3215)

#search narratives for bike or bicycle
ebike1<-grep(c("bike*| bicycle*"), df_19$narrative_1, ignore.case = TRUE)
length(ebike1) 
df_19$ebike1<-0
df_19$ebike1[ebike1]<-1

#search narratives for elec or batt or motor
ebike2<-grep(c("elect*|batt*|motor*"), df_19$narrative_1, ignore.case = TRUE)
length(ebike2) 
df_19$ebike2<-0
df_19$ebike2[ebike2]<-1

#combine so that have narratives that contain ebike 1 AND ebike2 conditions
df_19$ebike12<-0
df_19$ebike12[df_19$ebike1==1 & df_19$ebike2==1]<-1

df_19$ebike<-0
df_19$ebike[df_19$ebike12==1 & df_19$Codes5035.3215==1]<-1

# remove interim variables
df_19<-df_19[,!names(df_19) %in% c("ebike1", "ebike2", "ebike12")]

#BIKE 

# identify codes for bicycles (code 5040) and mountain bikes (code 5033)

df_19$bike1<-0
df_19$bike1[df_19$product_1==5040 | df_19$product_2==5040 | df_19$product_3==5040]<-1

df_19$bike2<-0
df_19$bike2[df_19$product_1==5033 | df_19$product_2==5033 | df_19$product_3==5033]<-1

df_19$bike<-0
df_19$bike[df_19$bike1==1 | df_19$bike2==1]<-1

# remove interim variables
df_19<-df_19[,!names(df_19) %in% c("bike1", "bike2","Code5042", "Codes5035.3215")]

```
## 2020-2022 micromobility variables

e-scooter, hoverboard, e-bike, bikes
```{r}

#note: all coded as 0=no, 1=yes

#POWERED SCOOTER (5022, new code)
# scooter code
df_20_21_22$Code5042<-0
df_20_21_22$Code5042[df_20_21_22$product_1==5022 | df_20_21_22$product_2==5022 | df_20_21_22$product_3==5022]<-1

# scooter indicator
scooter1<-grep(c("scooter"), df_20_21_22$narrative_1, ignore.case = TRUE)
length(scooter1) 
df_20_21_22$scooter1<-0
df_20_21_22$scooter1[scooter1]<-1

df_20_21_22$scooter<-0
df_20_21_22$scooter[df_20_21_22$scooter1==1 & df_20_21_22$Code5042==1]<-1

# remove interim variables
df_20_21_22<-df_20_21_22[,!names(df_20_21_22) %in% c("scooter1")]

#  powered skateboard (5025, new code for powered skateboard and hoverboard)

df_20_21_22$Code5025<-0
df_20_21_22$Code5025[df_20_21_22$product_1==5025 | df_20_21_22$product_2==5025 | df_20_21_22$product_3==5025]<-1


skateboard1<-grep(c("skate|skate board|skateboard"),df_20_21_22$narrative_1, ignore.case = TRUE)
length(skateboard1) 
df_20_21_22$skateboard1<-0
df_20_21_22$skateboard1[skateboard1]<-1

df_20_21_22$skateboard<-0
df_20_21_22$skateboard[df_20_21_22$skateboard1==1 & df_20_21_22$Code5025==1]<-1

# remove interim variables
df_20_21_22<-df_20_21_22[,!names(df_20_21_22) %in% c("skateboard1")]

#HOVERBOARD - 
# hoverboard variable use 5025
hover1<-grep(c("hoverboard|hooverboard|hover board|hoover board|hoverbaord|hooverbaord|hover boad|hoverboad,|hoverbrd"), df_20_21_22$narrative_1, ignore.case = TRUE)
length(hover1) 
df_20_21_22$hover1<-0
df_20_21_22$hover1[hover1]<-1

df_20_21_22$hover<-0
df_20_21_22$hover[df_20_21_22$hover1==1 & df_20_21_22$Code5025==1]<-1

# remove interim variables
df_20_21_22<-df_20_21_22[,!names(df_20_21_22) %in% c("hover1")]

#ELECTRIC BIKE

# powered bike code
df_20_21_22$Codes5035.3215<-0
df_20_21_22$Codes5035.3215[df_20_21_22$product_1==5035 | df_20_21_22$product_2==5035 | df_20_21_22$product_3==5035| df_20_21_22$product_1==3215 | df_20_21_22$product_2==3215 |df_20_21_22$product_3==3215]<-1


# E-BIKE (5035, 3215)

#search narratives for bike or bicycle
ebike1<-grep(c("bike*| bicycle*"), df_20_21_22$narrative_1, ignore.case = TRUE)
length(ebike1) 
df_20_21_22$ebike1<-0
df_20_21_22$ebike1[ebike1]<-1

#search narratives for elec or batt or motor
ebike2<-grep(c("elect*|batt*|motor*"), df_20_21_22$narrative_1, ignore.case = TRUE)
length(ebike2) 
df_20_21_22$ebike2<-0
df_20_21_22$ebike2[ebike2]<-1

#combine so that have narratives that contain ebike 1 AND ebike2 conditions
df_20_21_22$ebike12<-0
df_20_21_22$ebike12[df_20_21_22$ebike1==1 & df_20_21_22$ebike2==1]<-1


df_20_21_22$ebike<-0
df_20_21_22$ebike[df_20_21_22$ebike12==1 & df_20_21_22$Codes5035.3215==1]<-1

# remove interim variables
df_20_21_22<-df_20_21_22[,!names(df_20_21_22) %in% c("ebike1", "ebike2", "ebike12")]

#BIKE 

# identify codes for bicycles (code 5040) and mountain bikes (code 5033)

df_20_21_22$bike1<-0
df_20_21_22$bike1[df_20_21_22$product_1==5040 | df_20_21_22$product_2==5040 | df_20_21_22$product_3==5040]<-1

df_20_21_22$bike2<-0
df_20_21_22$bike2[df_20_21_22$product_1==5033 | df_20_21_22$product_2==5033 | df_20_21_22$product_3==5033]<-1

df_20_21_22$bike<-0
df_20_21_22$bike[df_20_21_22$bike1==1 | df_20_21_22$bike2==1]<-1

# remove interim variables
df_20_21_22<-df_20_21_22[,!names(df_20_21_22) %in% c("bike1", "bike2", "Code5042","Codes5035.3215")]


```

##stack 2019 and 2020-2023 data
```{r}

# Find common vars
common_vars <- intersect(names(df_19), names(df_20_21_22))

# Subset both dataframes to include only common variables
df1_subset <- df_19[, common_vars, drop = FALSE]
df2_subset <- df_20_21_22[, common_vars, drop = FALSE]


# Stack 2019 and 2020-2023 data
df_clean <- rbind(df1_subset, df2_subset)

names(df_clean)

# Duplicate issue (11/30/2023 update)
#decision to remove the duplicates from DF

# Remove skateboard
df_clean<- df_clean %>%
  select(-c("skateboard"))

# Note: use df_clean_with_dups to get total numbers of injuries in raw data
df_clean_with_dups <-df_clean

# Use df_clean for any of the multimode situations removed
#this is the code to remove those situatiosn
df_clean <- df_clean[
  !(
    df_clean$scooter == 1 & 
    (df_clean$bike == 1 | df_clean$ebike == 1 |  df_clean$hover == 1) |
    df_clean$bike == 1 & 
    (df_clean$scooter == 1 | df_clean$ebike == 1 | df_clean$hover == 1) |
    df_clean$ebike == 1 & 
    (df_clean$scooter == 1 | df_clean$bike == 1  | df_clean$hover == 1) |
    df_clean$hover == 1 & 
    (df_clean$scooter == 1 | df_clean$bike == 1 | df_clean$ebike == 1)
  ),
]


```

```{r}

#create long format df 
df_long <- df_clean %>%
  pivot_longer(cols = c(bike, ebike, scooter, hover), names_to = "mode", values_to = "value")

# Filter rows where the value is 1
df_filtered <- df_long %>%
  filter(value == 1) %>%
  select(-c(value))

```

this marks the end of the text-string search df cleaning and beginning of the analysis for the paper

#manuscript analysis

##read and clean chatgpt data
```{r}
gpt_v1 <- read.csv("data/chatgpt_data/gpt_v1_22NOV2023.csv") #low detail
gpt_v2<-read.csv("data/chatgpt_data/gpt_v2_07DEC2023.csv") #high detail
gpt_v2_retest<-read.csv("data/chatgpt_data/gpt_v2_retest_08DEC2023.csv") #high detail retest
gpt_v3_retest<-read.csv("data/chatgpt_data/gpt_v3_retest_12DEC2023.csv") #high detail retest
gpt_v4<-read.csv("data/chatgpt_data/gpt_v4_12DEC2023.csv") #intermediate detail
gpt_v5_retest<-read.csv("data/chatgpt_data/gpt_v5_retest_13DEC2023.csv") #high detail retest
gpt_v6_retest<-read.csv("data/chatgpt_data/gpt_v6_retest_14DEC2023.csv") #high detail retest


#merge text-string and chatgpt dataframes
df_gpt_v1<-merge(df_filtered, gpt_v1,
                       by = "cpsc_case_number") 

df_gpt_v2<-merge(df_filtered, gpt_v2,
                        by = "cpsc_case_number")  

df_gpt_v2_retest<-merge(df_filtered, gpt_v2_retest,
                        by = "cpsc_case_number")  

df_gpt_v3_retest<-merge(df_filtered, gpt_v3_retest,
                        by = "cpsc_case_number")  


df_gpt_v4<-merge(df_filtered, gpt_v4,
                        by = "cpsc_case_number") 


df_gpt_v5_retest<-merge(df_filtered, gpt_v5_retest,
                        by = "cpsc_case_number") 


df_gpt_v6_retest<-merge(df_filtered, gpt_v6_retest,
                        by = "cpsc_case_number")


df_gpt_v6_retest<-merge(df_filtered, gpt_v6_retest,
                        by = "cpsc_case_number")



#create new helmet_status_r variables and select variables of interest
df_gpt_v1<- df_gpt_v1 %>%
 mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)

df_gpt_v2<- df_gpt_v2 %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)

df_gpt_v2_retest<- df_gpt_v2_retest %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)


df_gpt_v3_retest<- df_gpt_v3_retest %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)

df_gpt_v4<- df_gpt_v4 %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)

df_gpt_v5_retest<- df_gpt_v5_retest %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)


df_gpt_v6_retest<- df_gpt_v6_retest %>%
  mutate(helmet_status_r = case_when(
    helmet_positive == 1 ~ "wearing helmet",
    helmet_negative == 1 ~ "not wearing helmet",
    helmet_unknown_all == 1 ~ "helmet not mentioned",
    TRUE ~ "None"
  )) %>%
  select(cpsc_case_number, helmet_status, helmet_status_r, narrative_1)


```

##Table 2. Agreement Between the LLM and Text String-Search for Helmet Status (N = 54 569)

###Table 2. low detail
```{r}
# make all categorical variables factors
df_gpt_v1 <- mutate_at(df_gpt_v1, vars("helmet_status"), as.factor)

  tbl_summary(data = df_gpt_v1,
    by = helmet_status,
                 include = c(helmet_status, helmet_status_r),
    statistic = list(all_categorical() ~ "{n}"),
    label = list(
                 helmet_status_r ~ "helmet status r code"),
    missing_text = "Not reported",
    digits = list(all_categorical() ~ c(0, 1)))
    
```


###Table 2. high detail
```{r}
# make all categorical variables factors
df_gpt_v2 <- mutate_at(df_gpt_v2, vars("helmet_status"), as.factor)

  tbl_summary(data = df_gpt_v2,
    by = helmet_status,
                 include = c(helmet_status, helmet_status_r),
        statistic = list(all_categorical() ~ "{n}"),
    label = list(
                 helmet_status_r ~ "helmet status r code"),
    missing_text = "Not reported",
    digits = list(all_categorical() ~ c(0, 1)))

```

###Table 2. intermediate detail
```{r}
# make all categorical variables factors
df_gpt_v4 <- mutate_at(df_gpt_v4, vars("helmet_status"), as.factor)

  tbl_summary(data = df_gpt_v4,
    by = helmet_status,
                 include = c(helmet_status, helmet_status_r),
        statistic = list(all_categorical() ~ "{n}"),
    label = list(
                 helmet_status_r ~ "helmet status r code"),
    missing_text = "Not reported",
    digits = list(all_categorical() ~ c(0, 1)))

```


### Table 2. low detail kappa
```{r}

# Set options to display three decimal places
options(digits = 3)

perc_agree_v1 <- df_gpt_v1 %>% select(helmet_status, helmet_status_r)

unique(perc_agree_v1$helmet_status_r)
#0=helmet not mentioned
#1=wearing helmet
#2=not wearing helmet


# Create numeric variable for helmet_status
perc_agree_v1 <- perc_agree_v1 %>%
  mutate(helmet_status_numeric = recode_factor(helmet_status,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet"))


perc_agree_v1$helmet_status_numeric <- as.numeric(perc_agree_v1$helmet_status_numeric)


# Create numeric variable for helmet_status_r
perc_agree_v1 <- perc_agree_v1 %>%
  mutate(helmet_status_r_numeric = recode_factor(helmet_status_r,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet")) 

perc_agree_v1$helmet_status_r_numeric <- as.numeric(perc_agree_v1$helmet_status_r_numeric)

perc_agree_v1 <- perc_agree_v1 %>% select(helmet_status_numeric, helmet_status_r_numeric)


agree(perc_agree_v1) #doesnt give CI
kappa2(perc_agree_v1) #doesnt give CI
cohen.kappa(perc_agree_v1) #gives CI


# Calculate agreement and Kappa
agree_result <- agree(perc_agree_v1)
kappa_result <- kappa2(perc_agree_v1)
cohen_kappa_result <- cohen.kappa(perc_agree_v1)

# Extract and round numeric values
agree_value <- round(agree_result$value, 3)
kappa_value <- round(kappa_result$value, 4)

# Print the rounded results
cat("Agreement:", as.character(agree_value), "\n")

# To get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

# To get CI to 3 decimal places
str(cohen_kappa_result)


```

### Table 2. high detail kappa
```{r}
perc_agree_v2 <- df_gpt_v2 %>% select(helmet_status, helmet_status_r) #rerun with df_gpt_v2_retest for retest results

unique(perc_agree_v2$helmet_status_r)
#0=helmet not mentioned
#1=wearing helmet
#2=not wearing helmet


# Create numeric variable for helmet_status
perc_agree_v2 <- perc_agree_v2 %>%
  mutate(helmet_status_numeric = recode_factor(helmet_status,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet"))


perc_agree_v2$helmet_status_numeric <- as.numeric(perc_agree_v2$helmet_status_numeric)


# Create numeric variable for helmet_status_r
perc_agree_v2<- perc_agree_v2 %>%
  mutate(helmet_status_r_numeric = recode_factor(helmet_status_r,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet")) 

perc_agree_v2$helmet_status_r_numeric <- as.numeric(perc_agree_v2$helmet_status_r_numeric)

perc_agree_v2 <- perc_agree_v2 %>% select(helmet_status_numeric, helmet_status_r_numeric)


agree(perc_agree_v2) #doesnt give CI
kappa2(perc_agree_v2) #doesnt give CI

cohen.kappa(perc_agree_v2)



# Calculate agreement and Kappa
agree_result <- agree(perc_agree_v2)
kappa_result <- kappa2(perc_agree_v2)
cohen_kappa_result <- cohen.kappa(perc_agree_v2)

# Extract and round numeric values
agree_value <- round(agree_result$value, 3)
kappa_value <- round(kappa_result$value, 4)

# Print the rounded results
cat("Agreement:", as.character(agree_value), "\n")

# To get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

# To get CI to 3 decimal places
str(cohen_kappa_result)
```


### Table 2. intermediate detail kappa
```{r}
perc_agree_v3 <- df_gpt_v4 %>% select(helmet_status, helmet_status_r) #rerun with df_gpt_v2_retest for retest results

unique(perc_agree_v3$helmet_status_r)
#0=helmet not mentioned
#1=wearing helmet
#2=not wearing helmet


# Create numeric variable for helmet_status
perc_agree_v3 <- perc_agree_v3 %>%
  mutate(helmet_status_numeric = recode_factor(helmet_status,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet"))


perc_agree_v3$helmet_status_numeric <- as.numeric(perc_agree_v3$helmet_status_numeric)


# Create numeric variable for helmet_status_r
perc_agree_v3<- perc_agree_v3 %>%
  mutate(helmet_status_r_numeric = recode_factor(helmet_status_r,
                                               "0" = "helmet not mentioned",
                                               "2" = "not wearing helmet",
                                               "1" = "wearing helmet")) 

perc_agree_v3$helmet_status_r_numeric <- as.numeric(perc_agree_v3$helmet_status_r_numeric)

perc_agree_v3<- perc_agree_v3 %>% select(helmet_status_numeric, helmet_status_r_numeric)


agree(perc_agree_v3) #doesnt give CI
kappa2(perc_agree_v3) #doesnt give CI

cohen.kappa(perc_agree_v3)


# Calculate agreement and Kappa
agree_result <- agree(perc_agree_v3)
kappa_result <- kappa2(perc_agree_v3)
cohen_kappa_result <- cohen.kappa(perc_agree_v3)

# Extract and round numeric values
agree_value <- round(agree_result$value, 3)
kappa_value <- round(kappa_result$value, 4)

# Print the rounded results
cat("Agreement:", as.character(agree_value), "\n")

# To get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

# To get CI to 3 decimal places
str(cohen_kappa_result)
```


##test retest reliablity 5 rounds

prep data
```{r}
# Create a list of your data frames
dfs <- list(df_gpt_v2, df_gpt_v2_retest, df_gpt_v3_retest, df_gpt_v5_retest, df_gpt_v6_retest)

# Create a vector of dates corresponding to each data frame
dates <- c("12072023", "12082023", "12122023", "12132023", "12142023")

# Function to process each data frame
process_dataframe <- function(df, date) {
  # Drop 'helmet_status_r' column if it exists
  if ("helmet_status_r" %in% colnames(df)) {
    df <- df %>% select(-helmet_status_r)
  }
  
  # Add a new column with the date
  df$date_column <- date
  
  return(df)
}

# Process each data frame in the list
processed_dfs <- Map(process_dataframe, dfs, dates)

# Combine the processed data frames into a long data frame
long_df <- bind_rows(processed_dfs)


```


##Table 3. Test-Retest Reliability for the LLM for High- Detail 5 Prompt Across 5Sessions and Days.
```{r}
table(long_df$helmet_status, long_df$date_column)
```


###kappa fleiss 
```{r}
variables <- list("helmet_status") 

kappa_fleiss <- function(subscale) {
  result <- long_df %>%
    select(all_of(subscale), date_column, cpsc_case_number) %>% 
    pivot_wider(names_from = date_column, values_from = subscale) %>% # make data wide
    select("12072023", "12082023", "12122023", "12132023", "12142023") %>% # remove description
    kappam.fleiss()
  
  return(result)
}

# Use the lapply() function to apply kappa_fleiss to each variable.
results <- lapply(variables, kappa_fleiss)

# Print the 'results' list
print(results)

str(results)


# Extract and round numeric values
p_value <- round(results$value, 3)
kappa_value <- round(kappa_result$value, 4)

# Print the rounded results
cat("Agreement:", as.character(agree_value), "\n")

#to get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

#to get CI to 3 decimal places
str(cohen_kappa_result)

```

##ground truth analysis

notes:  we used the high detail prompt where gpt performed the best on dec 7, 2023. 

notes about data cleaning and prep:

cleaning code for ground truth was not included in this code but we can provide additional code if requested.

from the high detail prompt df and the helmet status coded column in the df we randomly selected:
200 unknown
100 helmet
100 unhelmeted

1)three coders reviewed 1/3 of the sample
2)all three coders randomly reviewed 1/4 of sample (100 records)


##read ground truth df merged with gpt/r dataframe 
(cleaning code not provided)
```{r}
df<-read.csv("data/ground_truth_df.csv")
```

### prep data for confusion matrix

pull out values from each col to create confusion matrix
```{r}

#human
helmet_status_values <- df$helmet_status_human
helmet_status_values <- str_trim(helmet_status_values)
helmet_status_values <- gsub("\\s+", "", helmet_status_values)
cat("c(", paste(helmet_status_values, collapse = ", "), ")\n")

#gpt
helmet_status_values <- df$helmet_status
helmet_status_values <- str_trim(helmet_status_values)
helmet_status_values <- gsub("\\s+", "", helmet_status_values)
cat("c(", paste(helmet_status_values, collapse = ", "), ")\n")


#r
helmet_status_values <- df$helmet_status_r
helmet_status_values <- str_trim(helmet_status_values)
helmet_status_values <- gsub("\\s+", "", helmet_status_values)
cat("c(", paste(helmet_status_values, collapse = ", "), ")\n")


```

```{r}

#For more than two classes, the sensitivity, specificity, positive predictive value and negative predictive value are calculated comparing each factor level to the remaining levels (i.e. a "one versus all" approach).

# True labels (human coded)
true_labels <- c( 2, 2, 2, 2, 0, 0, 1, 0, 0, 1, 2, 2, 2, 0, 2, 2, 0, 0, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 0, 1, 2, 0, 2, 2, 1, 2, 0, 0, 1, 2, 2, 1, 0, 2, 0, 2, 2, 1, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 2, 2, 2, 1, 1, 1, 1, 1, 2, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 1, 1, 2, 2, 2, 1, 1, 2, 1, 2, 2, 0, 1, 2, 1, 0, 2, 2, 0, 2, 0, 2, 2, 0, 1, 2, 2, 1, 0, 2, 0, 0, 2, 0, 2, 2, 0, 1, 0, 2, 2, 2, 1, 2, 2, 1, 2, 2, 2, 1, 0, 2, 2, 0, 1, 2, 2, 2, 2, 2, 1, 2, 2, 2, 1, 1, 2, 2, 2, 2, 1, 0, 2, 0, 2, 0, 0, 0, 1, 0, 2, 2, 1, 1, 2, 2, 0, 1, 2, 2, 0, 2, 1, 2, 2, 2, 0, 0, 1, 1, 0, 1, 1, 2, 2, 1, 2, 2, 0, 2, 0, 2, 0, 2, 0, 0, 2, 2, 2, 2, 2, 0, 0, 2, 0, 2, 2, 1, 0, 1, 2, 2, 0, 2, 1, 0, 1, 0, 2, 1, 2, 2, 0, 0, 2, 2, 2, 1, 2, 2, 1, 2, 1, 1, 0, 2, 0, 1, 1, 2, 2, 1, 2, 0, 1, 2, 2, 2, 0, 2, 0, 1, 0, 1, 1, 2, 2, 1, 2, 2, 1, 1, 1, 0, 2, 0, 0, 1, 2, 1, 0, 2, 1, 1, 0, 2, 1, 1, 2, 1, 1, 2, 1, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 2, 2, 2, 0, 2, 0, 2, 2, 2, 2, 0, 2, 1, 2, 2, 2, 1, 1, 2, 2, 0, 0, 2, 0, 2, 0, 2, 2, 0, 1, 0, 1, 1, 1, 0, 1, 2, 2, 0, 2, 0, 1, 0, 0, 0, 2, 0, 2, 1, 0, 2, 0, 2, 2, 0, 2, 2, 1, 2, 1, 2, 2, 2, 0, 1, 2, 1, 1, 0, 0, 1, 2, 0, 2, 2, 0, 1, 2, 0, 2, 2, 0, 2, 1, 2, 0, 2, 0, 1, 2, 2, 1, 1, 0, 2, 1, 0, 2, 2, 2, 1, 2, 2, 2 )



# Predictions from GPT
predictions_method1 <- c(2, 2, 2, 2, 0, 2, 1, 0, 0, 1, 2, 2, 2, 0, 2, 2, 0, 0, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 0, 1, 2, 0, 2, 2, 1, 2, 0, 0, 1, 2, 2, 1, 0, 2, 0, 2, 2, 1, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 2, 2, 2, 1, 1, 1, 1, 1, 2, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 1, 1, 2, 2, 2, 1, 1, 2, 1, 2, 2, 0, 1, 2, 1, 0, 2, 2, 0, 2, 0, 2, 2, 0, 1, 2, 2, 1, 0, 2, 0, 0, 2, 0, 2, 2, 0, 1, 0, 2, 2, 2, 1, 2, 2, 1, 2, 2, 2, 1, 0, 2, 2, 0, 1, 2, 2, 2, 2, 2, 1, 2, 2, 2, 1, 1, 2, 2, 2, 2, 1, 0, 2, 0, 2, 0, 0, 0, 1, 0, 2, 2, 1, 1, 2, 2, 0, 1, 2, 2, 0, 2, 1, 2, 2, 2, 0, 0, 1, 1, 0, 1, 1, 2, 2, 1, 2, 2, 0, 2, 0, 2, 0, 2, 0, 0, 2, 2, 2, 2, 2, 0, 0, 2, 0, 2, 2, 2, 0, 1, 2, 2, 0, 2, 1, 0, 1, 0, 2, 1, 2, 2, 0, 0, 2, 2, 2, 1, 2, 2, 1, 2, 1, 1, 0, 2, 0, 1, 2, 2, 2, 1, 2, 1, 1, 2, 2, 2, 0, 2, 0, 1, 0, 1, 1, 2, 2, 1, 2, 2, 1, 1, 1, 0, 2, 0, 0, 1, 2, 1, 0, 2, 1, 1, 0, 2, 1, 1, 2, 1, 1, 2, 1, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 0, 2, 2, 2, 2, 0, 2, 1, 2, 2, 2, 1, 1, 2, 2, 0, 0, 2, 0, 2, 0, 2, 2, 0, 1, 0, 1, 1, 1, 0, 1, 2, 2, 0, 2, 0, 1, 0, 1, 0, 2, 0, 2, 1, 0, 2, 0, 2, 2, 0, 2, 2, 1, 2, 1, 2, 2, 2, 0, 1, 2, 1, 1, 0, 0, 1, 2, 0, 2, 2, 0, 1, 2, 0, 2, 2, 0, 2, 1, 2, 0, 2, 0, 1, 2, 2, 1, 1, 0, 2, 1, 0, 2, 2, 2, 1, 2, 2, 2 )

# Predictions from R
predictions_method2 <- c( 2, 2, 2, 2, 0, 2, 1, 0, 0, 1, 2, 2, 2, 0, 2, 2, 0, 0, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 0, 1, 2, 0, 2, 2, 1, 2, 0, 0, 1, 2, 2, 1, 0, 2, 0, 2, 2, 1, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 2, 2, 2, 1, 1, 1, 1, 1, 2, 1, 0, 0, 0, 2, 2, 2, 2, 1, 2, 0, 1, 1, 2, 2, 2, 1, 1, 2, 1, 2, 2, 0, 1, 2, 1, 0, 2, 2, 0, 2, 0, 2, 2, 0, 1, 2, 2, 1, 0, 2, 0, 0, 2, 0, 2, 2, 0, 1, 0, 2, 2, 2, 1, 2, 2, 1, 2, 2, 2, 1, 0, 2, 2, 0, 1, 2, 2, 2, 2, 2, 1, 2, 2, 2, 1, 1, 2, 2, 2, 2, 1, 0, 2, 0, 2, 0, 0, 0, 1, 0, 2, 2, 1, 1, 2, 2, 0, 1, 2, 2, 0, 2, 1, 2, 2, 2, 0, 0, 1, 1, 0, 1, 1, 2, 2, 1, 2, 2, 0, 2, 0, 2, 0, 2, 0, 0, 2, 2, 2, 2, 2, 0, 0, 2, 0, 2, 2, 2, 0, 1, 2, 2, 0, 2, 1, 0, 1, 0, 2, 1, 2, 2, 0, 0, 2, 2, 2, 1, 2, 2, 1, 2, 1, 1, 0, 2, 0, 1, 2, 2, 2, 1, 2, 1, 1, 2, 2, 2, 0, 2, 0, 1, 0, 1, 1, 2, 2, 1, 2, 2, 1, 1, 1, 0, 2, 0, 0, 1, 2, 1, 0, 2, 1, 1, 0, 2, 1, 1, 2, 1, 1, 2, 1, 2, 2, 2, 0, 2, 0, 1, 2, 2, 2, 2, 1, 2, 2, 2, 2, 2, 0, 2, 2, 2, 2, 0, 2, 1, 2, 2, 2, 1, 1, 2, 2, 0, 0, 2, 0, 2, 0, 2, 2, 0, 1, 0, 1, 1, 1, 0, 1, 2, 2, 0, 2, 0, 1, 0, 1, 0, 2, 0, 2, 1, 0, 2, 0, 2, 2, 0, 2, 2, 1, 2, 1, 2, 2, 2, 0, 1, 2, 1, 1, 0, 0, 1, 2, 0, 2, 2, 0, 1, 2, 0, 2, 2, 0, 2, 1, 2, 0, 2, 0, 1, 2, 2, 1, 1, 0, 2, 1, 0, 2, 2, 2, 1, 2, 2, 2 )

# Convert predictions_method1 and true_labels to factors with the same levels
true_labels <- factor(true_labels, levels = c("0", "1", "2"))
predictions_method1 <- factor(predictions_method1, levels = c("0", "1", "2"))
predictions_method2 <- factor(predictions_method2, levels = c("0", "1", "2"))


```

### Table 4. Performance of the LLM and Text String-Search Compared With Gold Criterion Standard (n = 400)
```{r}

# Create confusion matrices
# Recreate confusion matrices using confusionMatrix() function

# gpt
conf_matrix_method1 <- confusionMatrix(factor(predictions_method1), true_labels, positive = "1")
conf_matrix_method1 <- confusionMatrix(factor(predictions_method1), true_labels, positive = "1",  mode = "prec_recall")

# r
conf_matrix_method2 <- confusionMatrix(factor(predictions_method2), true_labels, positive = "1")
conf_matrix_method2 <- confusionMatrix(factor(predictions_method2), true_labels, positive = "1", mode = "prec_recall")


# kappa for gpt vs human
kappa <-df %>% select(helmet_status, helmet_status_human)
kappa_result<-kappa2(kappa) #doesnt give CI
kappa_value <- round(kappa_result$value, 4)


# to get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

cohen_kappa_result <- cohen.kappa(kappa)
# to get CI to 3 decimal places
str(cohen_kappa_result)


# kappa for r vs human
kappa <-df %>% select(helmet_status_r, helmet_status_human)
kappa_result<-kappa2(kappa) #doesnt give CI
kappa_value <- round(kappa_result$value, 4)


# to get kappa to 3 decimal places
cat("Kappa =", as.character(kappa_value), "\n")

cohen_kappa_result <- cohen.kappa(kappa)
# to get CI to 3 decimal places
str(cohen_kappa_result)


```

